Interval Probability for Fuzzy Quantum Theories

### 2016-06-10 prob.tex
Update citations...

In the remainder of the paper, we consider variations of quantum
probability spaces motivated by computation of numerical quantities in
a world with limited resources:
\begin{itemize}
\item Instead of the Hilbert space $\Hilb$ (constructed over the
  uncountable and uncomputable complex numbers $\mathbb{C}$), we will
  consider variants constructed over finite
  fields~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.
\item Instead of real-valued probability measures producing results in
  the uncountable and uncomputable interval $[0,1]$, we will consider
  finite set-valued probability measures~\cite{Artstein1972,PuriRalescu1983}.
\end{itemize}
We will then ask if it is possible to construct variants of quantum
probability spaces under these conditions. The main question is
related to the definition of probability measures: is it possible to
still define a probability measure as a function that depends on a
single state? Specifically,
\begin{itemize}
\item given a state $\ket{\psi}$, is there a probability measure
  mapping events to probabilities that only depends on $\ket{\psi}$?
  In the conventional quantum probability space, the answer is yes by
  the Born rule~\cite{Born1984,Mermin2007} and the map is given by:
  $E \mapsto \ip{\psi}{E\psi}$.
\item given a probability measure $\mu$
  mapping each event $E$
  to a probability, is there a \emph{unique} state $\psi$
  such that $\mu(E)
  =
  \ip{\psi}{E\psi}$? In the conventional case, the answer is yes by
  Gleason's
  theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
\end{itemize}

@@

For a given set of events $\events$, there are many possible probability
measures $\mu:\events\rightarrow[0,1]$. The Born rule~\cite{Born1984,Mermin2007,RiederSvozil2007},
a postulate of quantum mechanics, states that each pure quantum state
$\ket{\phi}$ induces a probability measure $\mu_{\phi}$ as follows:
\[
\mu_{\phi}(E)=\ip{\phi}{E\phi}
\]

@@

Vourdas observed that equation (\ref{eq:Kolmogorov}) is similar to
\[
\mathfrak{D}\left(S_{1},S_{2}\right)=\Pi\left(S_{1}\vee S_{2}\right)-\Pi\left(S_{1}\right)-\Pi\left(S_{2}\right)+\Pi\left(S_{1}\wedge S_{2}\right)\textrm{ ,}
\]
where $S_{1}$ and $S_{2}$ are arbitrary linear subspaces, and $\Pi$
maps a linear subspace to its projection operator. After apply the
Born rule on both sides, we get an equation about the probabilities
\begin{equation}
\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)=\Tr\left(\rho\Pi\left(S_{1}\vee S_{2}\right)\right)-\Tr\left(\rho\Pi\left(S_{1}\right)\right)-\Tr\left(\rho\Pi\left(S_{2}\right)\right)+\Tr\left(\rho\Pi\left(S_{1}\wedge S_{2}\right)\right)\textrm{ .}\label{eq:quantum-additivity}
\end{equation}
Given the density matrix~$\rho$, the sign of $\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)$
is then analog to the sign of $\delta\left(E_{1},E_{2}\right)$ since
the non-additivity operator~$\mathfrak{D}\left(S_{1},S_{2}\right)$
is non-zero when $S_{1}$ and $S_{2}$ does not commute so that $\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)$
is not always zero. Vourdas then considered the pair $\Tr\left(\rho\Pi\left(S_{1}\right)\right)$
and $\Tr\left(\rho\Pi\left(S_{2}\right)\right)$ are lower or upper
probabilities depending on the sign of $\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)$.
Although Vourdas' idea is clear, the approach does not exactly fit
the idea. First, when we measure a physical quantity, we consider
the eigenspaces of an observable which always commute. According to
Vourdas' definition, $\Tr\left(\rho\Pi\left(S_{1}\right)\right)$
is then always an Kolmogorov probability within one measurement process.
Second, if we adopt the frequency interpretation as in section \ref{subsec:Measuring-Probabilities:-Buffon},
and implement our one-qubit example in the Stern-Gerlach experiment.
Let $S_{1}=\left\{ \lambda\ket{0}\middle|\lambda\in\C\right\} $ and
$S_{2}=\left\{ \lambda\ket{\ps}\middle|\lambda\in\C\right\} $.

@@

Although it seems that we need an infinite long table to specify the
quantum probability measure~$\mu$, our $\mu$ is actually
given by a simple formula~$\melement{0}{E}$. In general, Born discovered
each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\tilde{\mu}_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows~\cite{Born1984,Mermin2007}:
\begin{equation}
\tilde{\mu}_{\psi}(E)=\frac{\melement{\psi}{E}}{\ip{\psi}{\psi}}\label{eq:Born}
\end{equation}
The Born rule satisfies the following properties:
\begin{itemize}
\item It can be extend to mixed states. Given a mixed state represented
by a density matrix $\rho=\sum_{j=1}^{N}q_{j}\frac{\proj{\psi_{j}}}{\ip{\psi_{j}}{\psi_{j}}}$,
where $\sum_{j=1}^{N}q_{j}=1$, i.e., $\Tr\left(\rho\right)=1$, then
the Born rule can be extended to $\rho$ by 
\begin{eqnarray}
\tilde{\mu}_{\rho}\left(E\right) & = & \Tr\left(\rho E\right)=\sum_{j=1}^{N}q_{j}\tilde{\mu}_{\Psi_{j}}\left(E\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}
Notice that $\left(\left\{ 1,\ldots,N\right\} ,2^{\left\{ 1,\ldots,N\right\} },\pmeas\left(J\right)=\sum_{j\in J}q_{j}\right)$
is a classical probability space. Therefore, when we discretize the
Hilbert space later, we may need to discretize this probability space
as well.
\item $\tilde{\mu}_{\rho}$ is a probability measure for all mixed state~$\rho$.
\item $\ip{\psi}{\phi}=0\Leftrightarrow\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=0$.
\item $\tilde{\mu}_{\psi}\left(E\right)=\tilde{\mu}_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}E\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
$\mathbf{U}$ is any unitary map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
\end{itemize}

>>>>

In the remainder of the paper, we consider variations of quantum
probability spaces motivated by computation of numerical quantities in
a world with limited resources:
\begin{itemize}
\item Instead of the Hilbert space $\Hilb$ (constructed over the
  uncountable and uncomputable complex numbers $\mathbb{C}$), we will
  consider variants constructed over finite
  fields~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.
\item Instead of real-valued probability measures producing results in
  the uncountable and uncomputable interval $[0,1]$, we will consider
  finite set-valued probability measures~\cite{Artstein1972,PuriRalescu1983}.
\end{itemize}
We will then ask if it is possible to construct variants of quantum
probability spaces under these conditions. The main question is
related to the definition of probability measures: is it possible to
still define a probability measure as a function that depends on a
single state? Specifically,
\begin{itemize}
\item given a state $\ket{\psi}$, is there a probability measure
  mapping events to probabilities that only depends on $\ket{\psi}$?
  In the conventional quantum probability space, the answer is yes by
  the Born rule~\cite{Born1983,Mermin2007} and the map is given by:
  $E \mapsto \ip{\psi}{E\psi}$.
\item given a probability measure $\mu$
  mapping each event $E$
  to a probability, is there a \emph{unique} state $\psi$
  such that $\mu(E)
  =
  \ip{\psi}{E\psi}$? In the conventional case, the answer is yes by
  Gleason's
  theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
\end{itemize}

@@

For a given set of events $\events$, there are many possible probability
measures $\mu:\events\rightarrow[0,1]$. The Born rule~\cite{Born1983,Mermin2007,RiederSvozil2007},
a postulate of quantum mechanics, states that each pure quantum state
$\ket{\phi}$ induces a probability measure $\mu_{\phi}$ as follows:
\[
\mu_{\phi}(E)=\ip{\phi}{E\phi}
\]

@@

Vourdas observed that equation (\ref{eq:Kolmogorov}) is similar to
\[
\mathfrak{D}\left(S_{1},S_{2}\right)=\Pi\left(S_{1}\vee S_{2}\right)-\Pi\left(S_{1}\right)-\Pi\left(S_{2}\right)+\Pi\left(S_{1}\wedge S_{2}\right)\textrm{ ,}
\]
where $S_{1}$ and $S_{2}$ are arbitrary linear subspaces, and $\Pi$
maps a linear subspace to its projection operator. After apply the
Born rule on both sides, we get an equation about the probabilities
\begin{equation}
\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)=\Tr\left(\rho\Pi\left(S_{1}\vee S_{2}\right)\right)-\Tr\left(\rho\Pi\left(S_{1}\right)\right)-\Tr\left(\rho\Pi\left(S_{2}\right)\right)+\Tr\left(\rho\Pi\left(S_{1}\wedge S_{2}\right)\right)\textrm{ .}\label{eq:quantum-additivity}
\end{equation}
Given the density matrix~$\rho$, the sign of $\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)$
is then analog to the sign of $\delta\left(E_{1},E_{2}\right)$ since
the non-additivity operator~$\mathfrak{D}\left(S_{1},S_{2}\right)$
is non-zero when $S_{1}$ and $S_{2}$ does not commute so that $\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)$
is not always zero. Vourdas then considered the pair $\Tr\left(\rho\Pi\left(S_{1}\right)\right)$
and $\Tr\left(\rho\Pi\left(S_{2}\right)\right)$ are lower or upper
probabilities depending on the sign of $\Tr\left(\rho\mathfrak{D}\left(S_{1},S_{2}\right)\right)$.
Although Vourdas' idea is clear, the approach does not exactly fit
the idea. First, when we measure a physical quantity, we consider
the eigenspaces of an observable which always commute. According to
Vourdas' definition, $\Tr\left(\rho\Pi\left(S_{1}\right)\right)$
is then always an Kolmogorov probability within one measurement process.
Second, if we adopt the frequency interpretation as in section \ref{subsec:Measuring-Probabilities:-Buffon},
and implement our one-qubit example in the Stern-Gerlach experiment~\cite{Stern1988,peres1995quantum,544199,Griffiths2003}.
Let $S_{1}=\left\{ \lambda\ket{0}\middle|\lambda\in\C\right\} $ and
$S_{2}=\left\{ \lambda\ket{\ps}\middle|\lambda\in\C\right\} $.

@@

Although it seems that we need an infinite long table to specify the
quantum probability measure~$\mu$, our $\mu$ is actually
given by a simple formula~$\melement{0}{E}$. In general, Born discovered
each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\tilde{\mu}_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows~\cite{Born1983,Mermin2007}:
\begin{equation}
\tilde{\mu}_{\psi}(E)=\frac{\melement{\psi}{E}}{\ip{\psi}{\psi}}\label{eq:Born}
\end{equation}
The Born rule satisfies the following properties:
\begin{itemize}
\item It can be extend to mixed states. Given a mixed state represented
by a density matrix $\rho=\sum_{j=1}^{N}q_{j}\frac{\proj{\psi_{j}}}{\ip{\psi_{j}}{\psi_{j}}}$,
where $\sum_{j=1}^{N}q_{j}=1$, i.e., $\Tr\left(\rho\right)=1$, then
the Born rule can be extended to $\rho$ by 
\begin{eqnarray}
\tilde{\mu}_{\rho}\left(E\right) & = & \Tr\left(\rho E\right)=\sum_{j=1}^{N}q_{j}\tilde{\mu}_{\Psi_{j}}\left(E\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}
Notice that $\left(\left\{ 1,\ldots,N\right\} ,2^{\left\{ 1,\ldots,N\right\} },\pmeas\left(J\right)=\sum_{j\in J}q_{j}\right)$
is a classical probability space. Therefore, when we discretize the
Hilbert space later, we may need to discretize this probability space
as well.
\item $\tilde{\mu}_{\rho}$ is a probability measure for all mixed state~$\rho$.
\item $\ip{\psi}{\phi}=0\Leftrightarrow\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=0$.
\item $\tilde{\mu}_{\psi}\left(E\right)=\tilde{\mu}_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}E\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
$\mathbf{U}$ is any unitary map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
\end{itemize}
